### 2019-06-14 mcmc_draws_2019-06-11/tables.Rmd
simulation changes

---
title: "Simulation Tables"
author: "Carter Allen"
date: "4/11/2019"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r}
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

```{r}
col_summarize <- function(MAT,dig = 2)
{
    k <- ncol(MAT)
    ret_vec <- rep(0,k)
    for(i in 1:k)
    {
        dat <- MAT[,i]
        est <- round(median(dat),dig)
        cri <- round(unname(quantile(dat,probs = c(0.025,0.975))),dig)
        res <- paste(est," (",cri[1],", ",cri[2],")",sep = "")
        ret_vec[i] <- res
    }
    return(ret_vec)
}
```

```{r}
beta_true_c1 <- c(beta.true.list[[1]]) %>% round(.,2)
beta_true_c2 <- c(beta.true.list[[2]]) %>% round(.,2)
beta_true_c3 <- c(beta.true.list[[3]]) %>% round(.,2)
beta_est_c1 <- col_summarize(BETA.list[[1]]) 
beta_est_c2 <- col_summarize(BETA.list[[2]]) 
beta_est_c3 <- col_summarize(BETA.list[[3]])

sig2_true_c1 <- c(sig2.true.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))] %>% round(.,2)
sig2_true_c2 <- c(sig2.true.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))] %>% round(.,2)
sig2_true_c3 <- c(sig2.true.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))] %>% round(.,2)
sig2_est_c1 <- col_summarize(SIGMA.list[[1]])[c(lower.tri(sig2.true.list[[1]],diag = TRUE))]
sig2_est_c2 <- col_summarize(SIGMA.list[[2]])[c(lower.tri(sig2.true.list[[2]],diag = TRUE))]
sig2_est_c3 <- col_summarize(SIGMA.list[[3]])[c(lower.tri(sig2.true.list[[3]],diag = TRUE))]

psi_true_c1 <- psi.true.list[[1]] %>% round(.,2)
psi_true_c2 <- psi.true.list[[2]] %>% round(.,2)
psi_true_c3 <- psi.true.list[[3]] %>% round(.,2)
psi_est_c1 <- col_summarize(PSI.list[[1]]) 
psi_est_c2 <- col_summarize(PSI.list[[2]]) 
psi_est_c3 <- col_summarize(PSI.list[[3]])

delta_true <- c(delta.true) %>% round(.,2)
delta_est <- col_summarize(DELTA)

pi_est <- col_summarize(PI)
pi_true <- round(pi.true,2)

model_component <- c("MVSN","Regression","","","","","","","",
                     "Multinom.","",
                     "Clustering")
params <- c("$\\beta_{0}$","$\\beta_{1}$","$\\beta_{2}$","$\\beta_{3}$",
            "$\\sigma_{11}$","$\\sigma_{12}$",
            "$\\sigma_{22}$",
            "$\\psi_{1}$","$\\psi_{2}$",
            "$\\delta_{11}$","$\\delta_{12}$",
            "$\\pi_l$")

true_c1 <- c(beta_true_c1,sig2_true_c1,psi_true_c1,delta_true,pi_true[1])
true_c2 <- c(beta_true_c2,sig2_true_c2,psi_true_c2,delta_true,pi_true[2])
true_c3 <- c(beta_true_c3,sig2_true_c3,psi_true_c3,delta_true,pi_true[3])
est_c1 <- c(beta_est_c1,sig2_est_c1,psi_est_c1,delta_est,pi_est[1])
est_c2 <- c(beta_est_c2,sig2_est_c2,psi_est_c2,delta_est,pi_est[2])
est_c3 <- c(beta_est_c3,sig2_est_c3,psi_est_c3,delta_est,pi_est[3])

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,
            true_c2,est_c2,
            true_c3,est_c3)
colnames(t1) <- c("Model Component",
                  "Parameter",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1500, k = 2, p = 2, h = 3, v = 1. 5000 iterations were run with a burn in of 1000. Missingness mechanism was MAR and P(miss) = 0") %>%
  kable_styling(font_size = 8) %>%
  add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
  group_rows(" ",1,4) %>%
  group_rows(" ",5,7) %>%
  group_rows(" ",8,10) %>%
  group_rows(" ",10,10) %>%
  group_rows(" ",12,12)
```

```{r,eval=FALSE}

kable(t1,booktabs = T,escape = F,caption = "Model results for simulated data with n = 500, k = 4, p = 2, h = 3, v = 3. 500 iterations were run with a burn in of 100. Missingness mechanism was MAR and P(miss) = 0.25.",
      row.names = NA) %>%
    kable_styling(font_size = 8) %>%
    add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
    group_rows(" ",1,4) %>%
    group_rows(" ",5,7) %>%
    group_rows(" ",8,10) %>%
    group_rows(" ",11,11) %>%
    group_rows(" ",12,12)
```