### 2019-07-09 mcmc_draws_2019-07-08/tables_2.Rmd

---
title: "Simulation Tables"
author:
date: 
output: 
  pdf_document: 
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r}
setwd("~/Documents/School/Summer_2019/Research/MVSN-FMM/mcmc_draws_2019-07-09")
load("BETA")
load("beta_true")
load("DELTA")
load("delta_true")
load("PI")
load("PSI")
load("psi_true")
load("SIGMA")
load("sig2_true")
load("pi_true")
load("Z")
```

```{r}
col_summarize <- function(MAT,dig = 2)
{
    k <- ncol(MAT)
    ret_vec <- rep(0,k)
    for(i in 1:k)
    {
        dat <- MAT[,i]
        est <- round(median(dat),dig)
        cri <- round(unname(quantile(dat,probs = c(0.025,0.975))),dig)
        res <- paste(est," (",cri[1],", ",cri[2],")",sep = "")
        ret_vec[i] <- res
    }
    return(ret_vec)
}
```

```{r}
L <- length(BETA.list)
n_sim <- nrow(BETA.list[[1]])
OMEGA.list <- SIGMA.list
ALPHA.list <- PSI.list
omega.true.list <- sig2.true.list
alpha.true.list <- psi.true.list
for(l in 1:L)
{
  SIGMA.l <- SIGMA.list[[l]]
  PSI.l <- PSI.list[[l]]
  OMEGA.l <- OMEGA.list[[l]]
  ALPHA.l <- ALPHA.list[[l]]
  for(i in 1:n_sim)
  {
    sig2.l.i <- matrix(SIGMA.l[i,],nrow = 4,ncol = 4,byrow = TRUE)
    psi.l.i <- PSI.l[i,]
    omega.l.i <- sig2.l.i + outer(psi.l.i,psi.l.i)
    alpha.l.i <- c(1/sqrt(1-t(psi.l.i) %*% solve(omega.l.i) %*% psi.l.i)) * (diag(sqrt(diag(omega.l.i))) %*% solve(omega.l.i) %*% psi.l.i)

    OMEGA.l[i,] <- c(omega.l.i)
    ALPHA.l[i,] <- c(alpha.l.i)
  }
  OMEGA.list[[l]] <- OMEGA.l
  ALPHA.list[[l]] <- ALPHA.l

  sig2_true.l <- sig2.true.list[[l]]
  psi_true.l <- psi.true.list[[l]]
  omega_true.l <- sig2_true.l + outer(psi_true.l,psi_true.l)
  alpha_true.l <- c(1/sqrt(1-t(psi_true.l) %*% solve(omega_true.l) %*% psi_true.l)) * (diag(sqrt(diag(omega_true.l))) %*% solve(omega_true.l) %*% psi_true.l)
  omega.true.list[[l]] <- omega_true.l
  alpha.true.list[[l]] <- alpha_true.l
}
```

```{r}
beta_true_c1 <- c(beta.true.list[[1]]) %>% round(.,2)
beta_true_c2 <- c(beta.true.list[[2]]) %>% round(.,2)
beta_true_c3 <- c(beta.true.list[[3]]) %>% round(.,2)
beta_est_c1 <- col_summarize(BETA.list[[1]]) 
beta_est_c2 <- col_summarize(BETA.list[[2]]) 
beta_est_c3 <- col_summarize(BETA.list[[3]])

omega_true_c1 <- c(omega.true.list[[1]])[c(lower.tri(omega.true.list[[1]],diag = TRUE))] %>% round(.,2)
omega_true_c2 <- c(omega.true.list[[2]])[c(lower.tri(omega.true.list[[2]],diag = TRUE))] %>% round(.,2)
omega_true_c3 <- c(omega.true.list[[3]])[c(lower.tri(omega.true.list[[3]],diag = TRUE))] %>% round(.,2)
omega_est_c1 <- col_summarize(OMEGA.list[[1]])[c(lower.tri(omega.true.list[[1]],diag = TRUE))]
omega_est_c2 <- col_summarize(OMEGA.list[[2]])[c(lower.tri(omega.true.list[[2]],diag = TRUE))]
omega_est_c3 <- col_summarize(OMEGA.list[[3]])[c(lower.tri(omega.true.list[[3]],diag = TRUE))]

alpha_true_c1 <- alpha.true.list[[1]] %>% round(.,2)
alpha_true_c2 <- alpha.true.list[[2]] %>% round(.,2)
alpha_true_c3 <- alpha.true.list[[3]] %>% round(.,2)
alpha_est_c1 <- col_summarize(ALPHA.list[[1]]) 
alpha_est_c2 <- col_summarize(ALPHA.list[[2]]) 
alpha_est_c3 <- col_summarize(ALPHA.list[[3]])

delta_true <- c(delta.true) %>% round(.,2)
delta_est <- col_summarize(DELTA)

pi_est <- col_summarize(PI)
pi_true <- round(pi.true,2)

model_component <- c("MVSN","Regression","","","","","","","","","","","","","","","","","","","","",
                     "Multinom.","","","",
                     "Clustering")
params <- c("$\\beta_{11}$","$\\beta_{21}$","$\\beta_{31}$","$\\beta_{41}$",
            "$\\beta_{12}$","$\\beta_{22}$","$\\beta_{32}$","$\\beta_{42}$",
            "$\\Omega_{11}$","$\\Omega_{12}$","$\\Omega_{13}$","$\\Omega_{14}$",
            "$\\Omega_{22}$","$\\Omega_{23}$","$\\Omega_{24}$",
            "$\\Omega_{33}$","$\\Omega_{34}$",
            "$\\Omega_{44}$",
            "$\\alpha_{1}$","$\\alpha_{2}$", "$\\alpha_{3}$","$\\alpha_{4}$",
            "$\\delta_{11}$","$\\delta_{12}$",
            "$\\delta_{21}$","$\\delta_{22}$",
            "$\\pi_l$")

true_c1 <- c(beta_true_c1,omega_true_c1,alpha_true_c1,delta_true,pi_true[1])
true_c2 <- c(beta_true_c2,omega_true_c2,alpha_true_c2,delta_true,pi_true[2])
true_c3 <- c(beta_true_c3,omega_true_c3,alpha_true_c3,delta_true,pi_true[3])
est_c1 <- c(beta_est_c1,omega_est_c1,alpha_est_c1,delta_est,pi_est[1])
est_c2 <- c(beta_est_c2,omega_est_c2,alpha_est_c2,delta_est,pi_est[2])
est_c3 <- c(beta_est_c3,omega_est_c3,alpha_est_c3,delta_est,pi_est[3])

t1 <- cbind(model_component,
            params,
            true_c1,est_c1,
            true_c2,est_c2,
            true_c3,est_c3)
colnames(t1) <- c("Model Component",
                  "Parameter",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)",
                  "True","Est. (95\\% CrI)")
```

```{r}
rownames(t1) <- NULL
kable(t1, booktabs = T, escape = F,caption = "Model results for simulated data with n = 1000, J = 4, p = 2, K = 3, r = 2. 1,000 iterations were run with a burn in of 0. Missingness mechanism was MAR and P(miss) = 0") %>%
  kable_styling(font_size = 8) %>%
  add_header_above(c(" " = 2,"Class 1" = 2, "Class 2" = 2, "Class 3" = 2)) %>%
  group_rows(" ",1,8) %>%
  group_rows(" ",9,18) %>%
  group_rows(" ",19,22) %>%
  group_rows(" ",23,26) %>%
  group_rows(" ",27,27)
```